Analyse gender in music psychology from journal publication data

### 2025-12-12 manuscript/_figure1_presentation.qmd
First submission

fig1 <- ggplot(annual2,aes(x=YearRange,y=Odds_ratio,color=Type,group=Type,shape=separator))+
  #facet_wrap(~TypeGroup,nrow = 2)+
  geom_hline(yintercept=1, linetype="dashed", color = "black") +
  geom_point(position = position_dodge(width = 0.4),show.legend=TRUE,size=3)+
  geom_errorbar(aes(ymin=CI_low,ymax=CI_high),width=0.2,position = position_dodge(width = 0.4))+
  geom_line(data=annual3,aes(x=YearRange,y=Odds_ratio,color=Type,group=Type,shape=separator),alpha = 0.75,position = position_dodge(width = 0.4))+
  scale_shape_manual(name="Authorship\ntype",values = c(22, 16),guide = "none") +
  scale_color_manual(name="Authorship\ntype",values=custom_palette) +
  xlab('Year range')+
  ylab('Female Authorship Odds Ratio (95%CI)')+
  theme_classic(base_size = 15)
print(fig1)

>>>>

fig1 <- ggplot(annual2,aes(x=YearRange,y=Odds_ratio,color=Type,group=Type,shape=separator))+
  #facet_wrap(~TypeGroup,nrow = 2)+
  geom_hline(yintercept=1, linetype="dashed", color = "grey75") +
  geom_point(position = position_dodge(width = 0.4),show.legend=TRUE,size=3)+
  geom_errorbar(aes(ymin=CI_low,ymax=CI_high),width=0.2,position = position_dodge(width = 0.4))+
  geom_line(data=annual3,aes(x=YearRange,y=Odds_ratio,color=Type,group=Type,shape=separator),alpha = 0.75,position = position_dodge(width = 0.4))+
  scale_shape_manual(name="Authorship\ntype",values = c(22, 16),guide = "none") +
  scale_color_manual(name="Authorship\ntype",values=custom_palette) +
  xlab('Year range')+
  ylab('Female Author Odds Ratio (95%CI)')+
  theme_classic(base_size = 15)
print(fig1)

# apply dark theme
fig1 <- fig1 + dark_theme_bw(base_size = 16) + theme(legend.position="top")

# make transparent

fig1<-fig1 + theme(
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  legend.background = element_rect(fill='transparent',color=NA),
  legend.box.background = element_rect(fill='transparent',color=NA)
)
fig1

ggsave(fig1,filename = "figure1_authorship_oddsratio_dark.png",width = 7.8,height = 5,dpi = 300)

### 2025-12-12 manuscript/_figure2_presentation.qmd
First submission

custom_palette <- brewer.pal(n = 4, name = "Set1")
custom_palette<-c(custom_palette[4],custom_palette[1:3])
fig2 <- ggplot(author_N,aes(x=author_N_cropped,y=Odds_ratio,color=Type))+
  geom_hline(yintercept=1, linetype="dashed", color = "black") +
  geom_point(position = position_dodge(width = 0.4),shape=15,size=3)+
  geom_errorbar(aes(ymin=CI_low,ymax=CI_high),width=0.2,position = position_dodge(width = 0.4))+
  scale_color_manual(name="Authorship\ntype",values = custom_palette) +
  xlab('Number of authors/article')+
  ylab('Female Authorship Odds Ratio')+
  theme_classic(base_size = 15)
#print(fig2)

s <- summarise(group_by(df, BIBTEXKEY), n = n())
S<-s
S$n[S$n > 9]<-10
fig2a <- ggplot(S,aes(x=n))+
  geom_histogram(binwidth=1,fill='grey90',color='black')+
  theme_classic(base_size = 15)+
  labs(x='Number of authors/article',y='Count of articles')+
  scale_x_continuous(breaks=seq(1,10,1),expand = c(0.001,0.001),limits = c(0,11),
                     labels = c(seq(1,9,1),"10+"))+
  scale_y_continuous(expand = c(0.001,0.001))
#print(g)

G<-cowplot::plot_grid(fig2a,fig2,nrow=1,labels=c('A','B'),rel_widths = c(0.4,1))
print(G)

>>>>

custom_palette <- brewer.pal(n = 4, name = "Set1")
custom_palette<-c(custom_palette[4],custom_palette[1:3])

fig2b <- ggplot(author_N,aes(x=author_N_cropped,y=Odds_ratio,color=Type))+
  geom_hline(yintercept=1, linetype="dashed", color = "gray50") +
  geom_point(position = position_dodge(width = 0.4),shape=15,size=3)+
  geom_errorbar(aes(ymin=CI_low,ymax=CI_high),width=0.2,position = position_dodge(width = 0.4))+
  scale_color_manual(name="Authorship\ntype",values = custom_palette) +
  xlab('Number of authors/article')+
  ylab('Female Authorship Odds Ratio')+
  theme_classic(base_size = 15)
#print(fig2)
library(ggdark)
fig2b <- fig2b + dark_theme_bw(base_size = 16) + theme(legend.position="top")

# make transparent
fig2b<-fig2b + theme(
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  legend.background = element_rect(fill='transparent',color=NA),
  legend.box.background = element_rect(fill='transparent',color=NA)
)

s <- summarise(group_by(df, BIBTEXKEY), n = n())
S<-s
S$n[S$n > 9]<-10
fig2a <- ggplot(S,aes(x=n))+
  geom_histogram(binwidth=1,fill='gold1',color='black')+
  theme_classic(base_size = 15)+
  labs(x='Number of authors/article',y='Count of articles')+
  scale_x_continuous(breaks=seq(1,10,1),expand = c(0.001,0.001),limits = c(0,11),
                     labels = c(seq(1,9,1),"10+"))+
  scale_y_continuous(expand = c(0.001,0.001))
#print(g)

# apply dark theme
fig2a <- fig2a + dark_theme_bw(base_size = 16) + theme(legend.position="top")

# make transparent
fig2a<-fig2a + theme(
  panel.background = element_rect(fill='transparent'),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  legend.background = element_rect(fill='transparent',color=NA),
  legend.box.background = element_rect(fill='transparent',color=NA)
)
fig2a

G<-cowplot::plot_grid(fig2a,fig2b,nrow=1,labels=c('A','B'),label_colour = "white",rel_widths = c(0.6,1))
print(G)
ggsave("figure2_authorship_by_number_of_authors_dark.png",
  G,
  width = 8.1,
  height = 4,
  dpi = 300
)
